Meaningful data sharing

### 2020-09-23 new-paper.Rmd
add some comments from meeting, finish draft of avenues section

@@

Make raw and analysis ready data available. Raw data is the first format of data provided before any cleaning has taken place (e.g., data in some binary or proprietary format). It should be placed in its own folder. Data used in analysis should be provided, especially if computationally intense to derive. It should be in its own folder, and to maximise interoperability, recorded in a plain text (e.g., csv, tsv), not a binary format (e.g., .rds, .sav, .xlsx). Ideally data should be in "Tidy Data" form, where variables are in columns, observations in rows, and each cell has only one value [@Wickham2014].

@@

Share code used to transform data from its raw format into analysis ready form. This should be kept in the same place as raw data (Section \@ref(fig:dir-layout)). Ideally, this transformation should involve only scripted languages (e.g, R, Python, Matlab, etc), but any non-scripted data preparation steps should be recorded in a plain text file.

@@

(**note: still need to work on this**)

Data should be deposited into a relevant domain repository. For example, gbif [@gbif] for species data, or genbif or genbank for genetics data @GenBank. Information on other domain specific repositories can be found at https://www.nature.com/sdata/policies/repositories.

As discussed in \@ref(cite), archive data using a platform generating an accession number such as a DOI. For sharing data collected for research, the most relevant homes for general data are the research data repositories **zenodo** and **dryad**. In addition to minting DOIs, they also provide citation templates.

If using a research institute's library service to share data, ensure a DOI is provided so it can be accurately identified and cited.

Do not deposit data in local file storage that does not provide DOI or accession numbers. For example, [@Hubbell1999] was hosted on a university server and only accessible via a FTP server since 1982. It was not possible to easily access or cite the data without an explicit authorship agreement among other conditions. Only recently were the complete dataset deposited in Dryad without such conditions [@condit-2019].

Although it is fine to share data among collaborators using temporary cloud storage (like Dropbox, Google Drive), once the paper is public, the data needs to be hosted somewhere persistent. While this approach is useful while iterating on the paper, it does not scale to a larger readership and is highly prone to disappearing (lack of permissions, etc.). Just archive the data.

**mention bit about**

- https://www.the-scientist.com/news-opinion/top-spider-biologists-research-under-fire-67083
- https://www.nature.com/articles/d41586-020-00287-y

@@

**bit removed from this section**

There are different avenues to share data, and although the terms are sometimes used interchangeably, they are, in fact, different. These include: (1) Data sharing: Making data publicly available; (2) Data archiving: Depositing data somewhere permanent; (3) Data publishing: writing a paper to explain the archived data.

>>>>

> Someone can mention links to a Dockerfile fro the README, or a text file that describes versions of software used, date when things last ran successfully etc.
> Also include the equivalent of a requirements.txt
> The date of use, and versions of software should be recorded in these scripts, so users can in the future can replicate similar computing environment or computing conditions. A more ideal approach would be to clearly describe the computing environment (e.g., a Dockerfile).

@@

Make raw and analysis ready data available. Raw data is the first format of data provided before any cleaning has taken place (e.g., data in some binary or proprietary format). It should be placed in its own folder. Data used in analysis should be provided, especially if computationally intense to derive. It should be in its own folder, and to maximise interoperability, recorded in a plain text (e.g., csv, tsv) and if possible a binary format (e.g., .rds, .sav, .xlsx, Apache Parquet). The binary format allows for immediate reuse in the intended state, while the interopability of plain text versions remain stable over time.

Ideally data should be in "Tidy Data" form, where variables are in columns, observations in rows, and each cell has only one value [@Wickham2014].

@@

Share code used to transform data from its raw format into analysis ready form. This should be kept in the same place as raw data (Section \@ref(fig:dir-layout)). Ideally, this transformation should involve only scripted languages (e.g, R, Python, Matlab, etc), but any non-scripted data preparation steps should be recorded in a plain text file. Authors should run the scripts in a clean environment to make sure everything runs and that the code doesn’t rely on other local artifacts.

@@

Data should be deposited into a relevant domain repository that generates an accession number such as a DOI. For example, gbif [@gbif] for species data, or genbank for genetics data [@GenBank]. Information on other domain specific repositories can be found at https://www.nature.com/sdata/policies/repositories. Data not falling into domain repositories can be submitted to research data repositories **zenodo** and **dryad**. In addition to minting DOIs, they also provide citation templates.

If using a research institute's library service to share data, ensure a DOI is provided so it can be accurately identified and cited. Do not deposit data in local file storage that does not provide DOI or accession numbers. For example, [@Hubbell1999] was hosted on a university server and only accessible via a FTP server since 1982. It was not possible to easily access or cite the data without an explicit authorship agreement among other conditions. Only recently were the complete dataset deposited in Dryad without such conditions [@condit-2019].

Although it is fine to share data among collaborators using temporary cloud storage (like Dropbox, Google Drive), once the paper is public, the data needs to be hosted somewhere persistent. While this approach is useful while iterating on the paper, it does not scale to a larger readership and is highly prone to disappearing (lack of permissions, etc.). Just archive the data. Providing back up data is particularly relevant in light of the case in [@Schleunes2020; @Viglione2020], where data was suspected of being tampered. The use of raw data allowed for issues to be identified, the absence of raw data not being available in other publications created difficulty in verifying the validity of the data.

@@
